---
title: "Lab2"
author: "Cale Bible"
date: "2025-09-04"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
list.files()
```

## Task 2
```{r}
mpg <- read.csv("EPAGAS.csv")
head(mpg)
```
## Task 3
```{r}
mpgv <- mpg$MPG
z <- (mpgv-mean(mpgv))/sd(mpgv)
# To get potential outliers, we can plot it using a box plot.
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))  
mypch = ifelse(abs(z)>3, 20,
        ifelse(abs(z)>=2 &abs(z)<=3,21, 22))
library(lattice)
# Blue symbolizes possible outliers, while red symbolizes the actual outliers.
dotplot(mpg$MPG,col=mycol)
```

## Task 4

```{r}
library(ggplot2)
 ggplot(data = mpg, aes(x = "", y = MPG)) +
      geom_boxplot(fill="black", notch="true") +
      coord_flip() +
      labs(title="MPG Distribution") +
      theme_minimal()
```

Empirical Formula:
1.  + or - one standard deviation from the mean: 68% of the data will lie there.
2. + or - two standard deviation from the mean: 95% of the data will lie there.
3. + or - three standard deviations from the mean: 99.7% of the data will lie there.

Using Chebyshev’s theorem predict the proportion of data within 2 standard deviations, we can conclude that in this data set, at least 75% of the data is between those parameters.

However, we can get a more exact percentage using r

```{r}
length(mpg$MPG[abs(z)<2])/length(mpg$MPG)
```

Comparing this output to  Chebyshev's rule, we notice that his rule does agree with the data! R told us 96%, while Chebyshev's rule told us at LEAST 75%. 

For the empirical rule, we know that:
1) Within 1 standard deviation, 68% of the data is contained
2) Within 2 standard deviations, 95% of the data is contained.
3) Within 3 standard deviations, 99.7% of the data is contained.

The empirical rule gives us a MUCH closer estimate than Chebyshev's rule in this case. It should be noted that the empiral rule can ONLY be applied to a standard bell curve.
